---
title: "RiboSeq_Report_Transcriptomic"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  pdf_document: default
  html_document: default
params:
  args: myarg
---
The following document is a report on the alignment of multiple Ribo-Seq datasets against the transcriptome. This alignment was intersected with the unique regions of the SplitORFs of NMD and RI transcripts, as calculated in the SplitORFs pipeline. Random regions of the 3' and 5' UTR with the same length distribution as the unique regions were created and intersected with the alignment as well. These serve as the null distribution, or background noise, as these regions should not be actively translated. From the null distribution a threshold is calculated as the upper limit of the 95% CI to classify whether the unique regions are found more often than expected by random chance.

```{r, echo=FALSE, libraries}
#library('knitr')
#library("dplyr")
#library("UpSetR")
source("helper_functions_Riboseq_report_genomic.R")
```

# Calculate the null distribution for NMD unique SplitORF regions

```{r}
args <- params[[1]]
path <- args[[1]]
# path <- "/Users/christina/Documents/Riboseq/Riboseq_genomic_own_ServeR_02_12_24"
outdir_NMD <- file.path(path, "NMD_genome")
outdir_RI <- file.path(path, "RI_genome")
```


```{r}

backgroundnmd <- calculate_background_threshold("NMD", path)
print_thresholds("NMD", backgroundnmd, path)
threshold=backgroundnmd
```


```{r}
#setwd("/Users/christina/Documents/ExtendedSplitORFPipeline-master_13_11_23/Output/BOWTIE_transcriptome/NMD_transcriptome")
#setwd(params[[1]][[2]])
list_nmddataframes <- get_intersect_files("NMD", path)
counts_and_names_list <- count_unique_regions_above_threhsold(list_nmddataframes, threshold)
print_unique_region_above_t(counts_and_names_list$counts_above_t_relative_length, list_nmddataframes)
```


The following table shows the top five unique regions with a relative count above the threshold (if available).
```{r}
get_top_5_unique_regions(counts_and_names_list$frames_processed, threshold)
```

The following plot shows how many unique regions are expressed in which dataset and how those overlap with the other data sets.

```{r, echo=FALSE,message=FALSE}
names(counts_and_names_list$unique_names) <- names(list_nmddataframes)
upsetlist <- counts_and_names_list$unique_names
upsetlist <- lapply(upsetlist, function(x) x$new_name)
upset(fromList(upsetlist), order.by = "freq", nsets = 12, point.size = 2.25, line.size = 1.5, set_size.show = TRUE,
      set_size.scale_max = 550, 
      mainbar.y.label = "Unique region Intersections", sets.x.label = "Matching unique regions per dataset",
      text.scale = c(1, 1, 0.9, 1, 1, 1))# c(1.5, 1.25, 0.9, 1.25, 1.25, 1.5)

```

The following table gives the number of unique regions with a count greater than or equal to 4:
```{r}
names(counts_and_names_list$frames_processed) <- names(list_nmddataframes)
count_unique_regions_get_count(counts_and_names_list$frames_processed, upsetlist, outdir_NMD)
```
Print the unique regions where at least 2 unique regions on the same transcript were found.
```{r}
Split_ORFs_validation(counts_and_names_list$frames_processed, upsetlist)
```




## Alignment against RI transcripts
```{r}
#setwd(params[[1]][[1]])
#setwd("/Users/christina/Documents/ExtendedSplitORFPipeline-master_13_11_23/Output/BOWTIE_transcriptome")
backgroundri <- calculate_background_threshold("RI", path)
print_thresholds("RI", backgroundri, path)
thresholdri=backgroundri
```

```{r}
#setwd("/Users/christina/Documents/ExtendedSplitORFPipeline-master_13_11_23/Output/BOWTIE_transcriptome/RI_transcriptome")
#setwd(params[[1]][[3]])
list_ridataframes <- get_intersect_files("RI", path)
counts_and_names_list <- count_unique_regions_above_threhsold(list_ridataframes, thresholdri)
print_unique_region_above_t(counts_and_names_list$counts_above_t_relative_length, list_ridataframes)
```


The following table shows the top five unique regions with a relative count above the threshold (if available).
```{r}
get_top_5_unique_regions(counts_and_names_list$frames_processed, thresholdri)
```




The following plot shows how many unique regions are expressed in which dataset and how those overlap with the other data sets.

```{r, echo=FALSE,message=FALSE}
names(counts_and_names_list$unique_names) <- names(list_ridataframes)
upsetlist <- counts_and_names_list$unique_names
upsetlist <- lapply(upsetlist, function(x) x$new_name)
upset(fromList(upsetlist), order.by = "freq", nsets = 12, point.size = 2.25, line.size = 1.5, set_size.show = TRUE,
      set_size.scale_max = 550, 
      mainbar.y.label = "Unique region Intersections", sets.x.label = "Matching unique regions per dataset",
      text.scale = c(1, 1, 0.9, 1, 1, 1))

```

The following table gives the number of unique regions with a count greater than or equal to 4:
```{r}
names(counts_and_names_list$frames_processed) <- names(list_ridataframes)
count_unique_regions_get_count(counts_and_names_list$frames_processed, upsetlist, outdir_RI)
```


Print the unique regions where at least 2 unique regions on the same transcript were found.
```{r}
Split_ORFs_validation(counts_and_names_list$frames_processed, upsetlist)
```